---
title: "Binary Process"
author: "Kevin O'Brien"
date: "11/29/2017"
output: html_document
params:
  code_dir: ../src/
  data_dir: ../data/
  model_dir: ../model/ 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(dplyr)
library(DiagrammeR)
```

## Rate $\theta$ for a Binary Process

Goal is to determine the posterior distribution for the rate $\theta$ after observing *k* successes in *n* trials.

#### Graphical Model
```{r}
source(paste(params$code_dir,"gm_binary_process.r",sep = ""))
gm_binary_process()
```

#### Data
```{r}
# Data
## TODO - switch to sim_data.r
data_file <- paste(params$data_dir, "data.r", sep = "")
data <- read_rdump(data_file) 
data
```


#### Stan Model
```{r}
model = c(
"data {
  int<lower=1> n; // at least one observation
  int<lower=0> k; // success count can not be negative
}

parameters {
  real<lower=0, upper=1> theta; // bounded by 0 and 1
}

model {
  //Prior Distribution for Rate Theta
  theta ~ beta(1, 1);
  
  // Observed Counts
  // k has a binomial distribution of n observations 
  // with theta variation
  k ~ binomial(n, theta);
}
")
cat(model)
```


#### Initialize Parameters
Set inital values by providing a list equal in length to the 
number of chains. The elements of this list should themselves 
be named lists, where each of these named lists has the name 
of a parameter and is used to specify the initial values for 
that parameter for the corresponding chain. If initial values 
are not provided, Stan provides random values.

In this example, start chain 1 at .1 and chain 2 at .9 since $\theta$ is bounded at 0 and 1.
```{r}
myinits <- list(
  list(theta=.1),
  list(theta=.9))
parameters <- c("theta")
```

#### Run Stan
Output is a ... named "samples".
```{r}
model_code <- paste(params$model_dir, "binomial_one_rate.stan", sep = "")

samples <- stan(file=model_code,
                data=data,
                init=myinits,
                pars=parameters,
                iter=20000,
                chains=2,
                thin=1
                # warmup = 100,  # Stands for burn-in; Default = iter/2
                # seed = 123  # Setting seed; Default is random seed
)
```

#### Quick Look
```{r}
print(samples)  # a rough summary
```

#### Detailed Look
```{r}
print(summary(samples))  # more detailed summary
```

#### A Graph
```{r}
plot(samples) 
```

