---
title: "Bernoulli Trials - Difference Between Two Processes"
output:
  html_document:
    fig_caption: yes
    fig_height: 3
    fig_width: 5
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
date: "12/06/2017"
params:
  code_dir: ../src/
  data_dir: ../data/
  model_dir: ../model/  
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(rstan)
library(bayesplot)
library(dplyr)
library(feather)
library(ggplot2)

```
## The Data

Independent observations with two possible outcomes on a trial. And some characteristic that allows separating the trials into to groups. Imagine an underlying process difference that might account for differences in success (e.g., coins from different mints have a different bias) and a single process $\delta$ to characterize the difference between them.

Simulate a data set __data__ with _k_ successes in _n_ trials and $\theta$ processes. Start with 10 trials and a success rate of 90% for one group and 10% for the other.

The __data__ contains a single process with observations and a descriptive outcome label. One observation per row.
```{r create_data, echo = FALSE, message = FALSE}

##### Data #####
data_name <- "two_rates"
obs_ct <- 10
rate_list <- c(0.9, 0.1)

# Simulate Data
source(paste(params$code_dir,"sim_data.r",sep = ""))
data_file <- paste(params$data_dir, data_name, sep = "")
sim_data(data_file, obs_ct, rate_list, "fixed")

# Read Data
data <- read_feather(paste(data_file, ".feather", sep = ""))
print(data)

```

Use __class(data)__ to learn the class of the r object.
```{r data_object}
class(data)
```

Use ggplot to generate a data graph showing successes, failures, and total observations.
```{r visualize_data}

# Visualize Data
ggplot(data, aes(x = process)) + geom_bar(aes(fill=outcome))
```

Stan needs summarized data ( _k_ successes in _n_ trials) in a list. Create the list *stan_data*.
```{r data_to_list, echo = FALSE, message = FALSE}
# Stan needs data in a list
success_k1 <- data %>%
  filter(outcome == 'success') %>%
  filter(process == 'theta_1')
success_k2 <- data %>%
  filter(outcome == 'success') %>%
  filter(process == 'theta_2')


stan_data <- list(
  n1 = nrow(data)/length(rate_list),
  k1 = nrow(success_k1),
  n2 = nrow(data)/length(rate_list),
  k2 = nrow(success_k2)
)
  print(stan_data)
```

Use __class(stan_data)__ to learn the class of the r object.
```{r stan_data_object}
class(stan_data)
```

## The Model

A graphical representation of the model shows:

 * a continuous, unobserved "process" parameter $\delta$
 * a continuous, unobserved "process" parameter $\theta$
 * a discrete, observable number of successes *k*
 * a discrete, observable number of observations *n*

The vector's arrow shows dependency. Successes are dependent on both the underlying process $\theta$ and the number of observations. The difference between $\theta$s is addressed by $\delta$.

The assumption for $\theta$ is that all possible rates are equally likely. The Beta distribution is set with 1 "success" and 1 "failure".

The assumption of *k* is that the outcomes are from a Binomial distribution with $\theta$ determining the rate for *n* observations.

```{r graphical_model, echo = FALSE, message = FALSE}
# Visualize Model
source(paste(params$code_dir,"gm_two_process.r",sep = ""))
gm_two_process()
```


The Stan model follows from the graphical model.
The data block names *n* and *k* and sets lower boundaries.
The parameters block names $\theta$ and sets upper and lower boundaries. Transformed parameters block names $\delta$ as the difference between the two $\theta$ processes.
The model block sets the prior distribution for $\theta$ and the model for *k1* and *k2*.
```{r stan_model, echo = FALSE, message = FALSE}
model = c(
"
data { 
  int<lower=1> n1; 
  int<lower=1> n2; 
  int<lower=0> k1;
  int<lower=0> k2;
} 
parameters {
  real<lower=0,upper=1> theta1;
  real<lower=0,upper=1> theta2;
} 
transformed parameters {
  real<lower=-1,upper=1> delta;
  // deterministic variable delta is the difference
  // between rates from two processes producing k1 and k2 outcomes
  delta = theta1 - theta2;
}
model {
  // Prior Distribution for Rate Theta
  theta1 ~ beta(1, 1);
  theta2 ~ beta(1, 1);
  // Observed Counts
  k1 ~ binomial(n1, theta1);
  k2 ~ binomial(n2, theta2);
}
")
cat(model)
```

## Run Stan
> "Run Stan with default settings and see what happens." 

(Andrew Gelman said this and I'm taking him at his word!)
```{r stan_set_up, message = FALSE, results = "hide", cache = TRUE}
model_code <- paste(params$model_dir, "binomial_two_rates.stan", sep = "")
# parameters to be monitored:  
parameters <- c("theta1", "theta2", "delta")

# Run Stan with defaults
samples_default <- stan(file=model_code,
                        data=stan_data,
                        refresh = 0)
```

Look at the object generated with **class(samples_default)**. The S4 object contains lots of information beyond the model results. 
```{r}
class(samples_default)
```

And, simply running basic commands against the Stan object will produce reports and graphs specific to the class.

A summary table:
```{r}
print(samples_default)
```

A plot of parameters
```{r}
plot(samples_default) 
```

And a better plot of $\theta$

